---
title: "Exploratory Data Analysis"
date: "July 24, 2018"
output:
  html_document:
    df_print: kable
    toc: yes
---

# Setup







```{r code = readLines('Setup.R')}
```


```{r setup, }


# knitr::opts_chunk$set(
#                       root.dir = normalizePath(".."),
#                       child.dir = normalizePath("..")
#                       )


getwd()

cool_1 = "steelblue"

```



# Import Datasets

```{r}



fracfocus <- read.csv("../data/fracfocus_registry.csv") %>%
              standardize_names() %>%
              rename(api='apinumber') %>%
              mutate( api=as.character(api)
                     ,api10 = as.character(api10)
                     ,jobstartdate=as.Date(jobstartdate, "%m/%d/%Y")
                     ,jobenddate=as.Date(jobenddate, "%m/%d/%Y")
                     ,totalbasewatervolume = as.double(totalbasewatervolume)
                     ,totalbasenonwatervolume = as.double(totalbasenonwatervolume)
                     ,percenthighadditive = as.double(percenthighadditive)
                     ,percenthfjob = as.double(percenthfjob)
                     ,massingredient = as.double(massingredient)
                     ,iswater = as.logical.factor(iswater)
                     ) %>%
              select( api
                     ,api10
                     ,wellname
                     ,operatorname
                     ,jobstartdate
                     ,jobenddate
                     ,latitude
                     ,longitude
                     ,totalbasewatervolume
                     ,totalbasenonwatervolume
                     ,statename
                     ,countyname
                     ,ingredientname
                     ,percenthighadditive
                     ,percenthfjob
                     ,massingredient
                     ,iswater
                     ) %>%
             as.tibble()

driftwood <- read.csv("data/deo_well_data.csv") %>%
             standardize_names() %>% 
             rename(oil='oilpknormperk6mo') %>%
             mutate(api = as.character(api)
                     , api10 = as.character(api10)
                     , perfll=as.integer(perfll)
                     , firstprod = as.date(firstprod, "%m/%d/%Y")
                     , oil = as.integer(oil)
                     , firstprod_year = as.integer(format(firstprod, '%Y'))
                     ) %>%
             as.tibble()

# driftwood %>% head()



```

```{r}

ff_distinct_api <- fracfocus %>%
    distinct(api)

wells_distinct_api14 <- driftwood %>%
    inner_join(fracfocus, by = c("api", "api")) %>%
    distinct(api)

wells_distinct_api10 <- driftwood %>%
    inner_join(fracfocus, by = c("api10", "api10")) %>%
    distinct(api10)


```





```{r}

# Aggregate sand and water
#-- 20/40, 40/70, 30/50, 100 Mesh


# Latitude = Important
# Longitude = Important
# Form_Avg = Very Important
# Sand = Very Important
# Water = Very Important
# Other chemicals / additives = ???


ff_summary <- fracfocus %>%
    filter('sand' %in% ingredientname
        | 'silica' %in% ingredientname
        | 'propp' %in% ingredientname
        | 'mesh' %in% ingredientname
        | 'brown' %in% ingredientname
        | 'white' %in% ingredientname
        | '30%50' %in% ingredientname
        | '40%70' %in% ingredientname
        | '30%50' %in% ingredientname
        | '100' %in% ingredientname
        ) %>%
    group_by(api10) %>%
    summarize(totalwater = max(totalbasewatervolume)
              , totalsand = sum(massingredient)
              , percenthfjob = sum(percenthfjob)
              # add additional summary variables here.
              )


ff_summary %>% head()

```



```{r}

well_data <- driftwood %>%
    inner_join(ff_summary, by = c("api10", "api10")) %>%
    mutate(lb_ft = totalsand / perfll
           , bbl_ft = totalwater / perfll #todo: totalwater bbl looks way too high.
           , gal_ft = bbl_ft / 42
           , distance = sqrt(gal_ft ^ 2 + lb_ft ^ 2) * (180 / pi) # linear_distance_from_origin (hypotenuse length)
           , angle = atan(gal_ft / lb_ft) * (180 / pi) #in degrees
           , deviation = abs(45 - angle) #deviation_from_45deg
           , weight = 500 # weighting_constant
           , frac_rating = distance-(deviation-weight) # weighted_distance_from_origin
           ) %>%
           select(-distance, -angle, -deviation, -weight)



well_data %>% head() 




```

<!-- #################### -->
<!-- ```{r} -->

<!-- well_data <- driftwood %>% -->
<!--     inner_join(ff_summary, by = c("api10", "api10")) %>% -->
<!--     mutate(lb_ft = totalsand / perfll -->
<!--            , bbl_ft = totalwater / perfll #todo: totalwater bbl looks way too high. -->
<!--            , gal_ft = bbl_ft / 42 -->
<!--            , distance = sqrt(gal_ft ^ 2 + lb_ft ^ 2) * (180 / pi) # linear_distance_from_origin (hypotenuse length) -->
<!--            , angle = atan(gal_ft / lb_ft) * (180 / pi) #in degrees -->
<!--            , deviation = abs(45 - angle) #deviation_from_45deg -->
<!--            , weight = 500 # weighting_constant -->
<!--            , frac_rating = distance-(deviation-weight) # weighted_distance_from_origin -->
<!--            ) %>% -->
<!--            select(-distance, -angle, -deviation, -weight) -->



<!-- well_data %>% head()  -->

<!-- ``` -->



<!-- ### investigate frac focus -->


<!-- ```{r} -->

<!-- ggplot((fracfocus %>% na.omit(.)), -->
<!--        aes(x = form_avg, y = oil)) + #todo: move to appendix -->
<!--        geom_boxplot(fill = cool_1) + -->
<!--        ggtitle("abv by state") + -->
<!--        xlab("states") + -->
<!--        ylab("alcohol by volume (%)") + -->
<!--        theme(text = element_text(size = 10), -->
<!--         axis.text.x = element_text(angle = 90, vjust = 0.5), -->
<!--         plot.title = element_text(hjust = 0.5, size = 16)) -->


<!-- ``` -->
<!-- #################### -->














##### descriptive analysis

```{r}
require(summarytools)
well_summary <- as.data.frame(summarytools::descr(well_data, transpose= true)) #%>% mutate_all(round, )#%>% mutate_at(vars(-mpg, -cyl), max)

well_summary %>% 
  rownames_to_column('attribute') %>%
  mutate_if(is.numeric, round) %>%
  column_to_rownames('attribute') %>%
  select(mean, median, min, max, std.dev, q1, q3, iqr, n.valid, pct.valid)


```

```{r}



```



##### additive concentrations vs productivity




##### avg frac size by year


```{r}

#frac job profile as frac lb_ft vs bbl_ft
ggplot((well_data), aes(x = lb_ft, y = bbl_ft)) +
    geom_point() 


```

```{r}

ggplot((well_data), aes(x = lb_ft, y = bbl_ft)) +
    geom_point()

```


```{r}
#todo: productivity at ~2000 bbl_oil/1000 ft looks suspicious

# distribution of wells by productivity
ggplot(well_data) +  
    geom_histogram(aes(x = oil), binwidth = 100) +
    theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, hjust = 1))

```



##### pk-6mo-oil by year


```{r}

#todo: oil productivity seems to increase with firstprod date.  needs further exploration.

# stat = identity seems to be doing a sum of productivity by year.  2018 is not complete yet.  normalize over number of wells to correct.

ggplot(well_data) +
    geom_histogram(aes(x = firstprod_year, y = oil), stat = "identity") +
    theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, hjust = 1))



```



##### is pknorm_6mo_oil normally distributed for each formation?


```{r fig.height=6, fig.width=12}

require(car)

# form_avg <- well_data$form_avg
# lb_ft <- well_data$lb_ft
# bbl_ft <- well_data$bbl_ft
# oil <- well_data$oil

features <- well_data %>% select(lb_ft, bbl_ft, form_avg, oil)

#cbind(lb_ft, bbl_ft, form_avg, oil)#cbind(brain, body, litter, gestation)
scatterplotmatrix(features
                  , smooth=false
                  , diagonal="histogram"
                  , col= "steelblue")  


```

```{r}
features <- features %>% 
  filter_all(any_vars(. > 0)) %>%
  mutate_at(.vars=vars(-form_avg), .funs=funs(round(.))) %>%
  #mutate_if(is.numeric, log(.)) %>%
  na.omit()

#cbind(log(lb_ft), log(bbl_ft), form_avg, log(oil))
      
      
scatterplotmatrix(features
                  , smooth=false
                  , diagonal="histogram"
                  , col= "steelblue")
```


```{r}


```


```{r}

simple.lm <- lm(oil ~ lb_ft + bbl_ft + form_avg)
moderate.lm <- lm(oil ~ lb_ft + bbl_ft + form_avg + lb_ft:bbl_ft)
```

```{r}
plot(simple.lm)
```
```{r}
plot(moderate.lm)
```

```{r}
anova(simple.lm, moderate.lm)
```

```{r}

# component residual plots

crplots(simple.lm)
```





1.	Research Question – Clearly stated

    - does well productivity increase with 'frac size'? how does it increase?  are there regional trends? trends by formation?
        - do any additives affect oil productivity? regional? trends by formation?

2.	Hypothesis 

 - Examples:
     - Group Differences:
         - addatives/chemicals
         - what specific additives are associated with increased producivity
            - H~0~: mu_oil_additive1 = 0
            - H~A~: mu_oil_additive1 <> 0 


     - Regression:
        - Explanatory: mu_oil_additive1
        - Response: Oil

     - Multiple Regression
        - Explanatory: sand, water
        - Response: Oil    

        - Explanatory: long, lat, form_avg
        - Response: Oil       


3.	A Code book - Matthew
        - Convert original data codebooks to markdown
        - Build codebooks for our analysis

4.	READme file - Quinton
        - Build Readme    
        - Code Commenting

5.	Exploratory Data - All
    a.	Please provide (in table format or similar), descriptive statistics on at least 7 variables (age, Income, etc.).  Create a simple histogram for two of them.  Comment on the shape of the distribution in your markdown.
    b.	Give the frequencies (in table format or similar) for Gender, Education, and Occupation.  They can be separate tables, if that’s your choice

6. Markdown/Code - Brock

    a.	One Linear regressions


    b.	One model of your choice (can be linear regression, thus 2 linear regressions total)
        - linear regression
        - multiple regression/anova

    c.	Results
 

    d.	Discussion


    e.	Conclusion


11. Presentation - All






























