---
title: "Measuring Well Productivity by Completion Design"
author: | 
        | Flaming  
        | Trevathan
        | Nixon  
        | Friedrich
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
    toc: true
    number_sections: true
    df_print: kable
    fontsize: 11pt
    fontfamily: palatino

date: "July 24, 2018"

---

<!-- ```{r code = readLines('Setup.R'), message=FALSE, warning=FALSE} -->
<!-- ``` -->

```{r setup, echo=FALSE, results="hide"}
# run setup files, create custom functions, and read in data for analysis
source("Setup.R")
source("Functions.R")
read_chunk('DataImport-FracFocus.R')
read_chunk('DataImport-WellFeatures.R')
read_chunk('Explore_Data.R')

```


# Executive Summary

FTNF Analytics is a boutique firm specializing in data analysis for the energy sector.  The FTNF team includes experts in geology, data visualization, statistics, and predictive modeling.  Our practice services producers globally with a special emphasis on those operating in the Southwest United States.

The Consortium of Texas Oil Producers (CTOP), has engaged FTNF to develop a study to better understand the potential for production in designated areas and help decide the best way to allocate their valuable resources.  Specifically, CTOP needs a way to predict potential energy production (and revenues generated).

FTNF is the ideal partner for this undertaking.  Our process includes enriching existing CTOP data to improve the ability to predict production outcomes.  We start with exploring the data, creating new data, and augmenting the data.  Once satisfied, we proceed to analyzing the data using contemporary statistical methods.  From there, we build predictive models to provide insight necessary to make drilling decisions.

CTOP companies no longer have to rely on hope as a strategy for finding productive wells.  With the help of FTNF Analytics, you can rely on sound science and quantifiable predictions.




# Frac Focus Data


## Import

```{r fracfocus_import, cache = TRUE}
```

## Count Locations and Wellbores

```{r fracfocus_count, cache = TRUE}
```

## Aggregate Proppant and Water by wellbore

```{r fracfocus_summary, cache = TRUE}
```

## Aggregates

```{r fracfocus_aggregates, cache = TRUE}
```

# Well Features and Characteristics

```{r wellfeatures_import, cache = TRUE}
```

## Count Locations and Wellbores

```{r wellfeatures_count, cache = TRUE}
```

## Aggregates

```{r fracfocus_aggregates, cache = TRUE}
```



# Data Exploration

## Add Calculated Columns

```{r exp_welldata_calcs}
```

## Clean data

```{r exp_welldata_calcs}
```

## Summarize Well Data 

```{r exp_summary}
```

```{r exp_boxplot_fracsize}
```

```{r exp_boxplot_fracsize}
```


### Frequencies

```{r exp_freq_vintage}
```


```{r exp_freq_status}
```



# Appendix:

```{r}
sessionInfo()
```