---
title: "Measuring Well Productivity by Completion Design"
author: | 
        | Flaming  
        | Trevathan
        | Nixon  
        | Friedrich
output:
  prettydoc::html_pretty:
    keep_md: yes
    theme: leonids
    highlight: github
    toc: true
    number_sections: true
    df_print: kable
    fontsize: 11pt
    fontfamily: palatino

date: "July 24, 2018"

---

<!-- ```{r code = readLines('Setup.R'), message=FALSE, warning=FALSE} -->
<!-- ``` -->

```{r setup, echo=FALSE, results="hide"}
# run setup files, create custom functions, and read in data for analysis
source("Setup.R")
source("Functions.R")
read_chunk('DataImport-FracFocus.R')
read_chunk('DataImport-WellFeatures.R')
read_chunk('Explore_Data.R')

```


# Executive Summary

FTNF Analytics is a boutique firm specializing in statistics and data analysis for the energy sector.  The FTNF team includes experts in geology, data visualization, statistics, and predictive modeling.  Our practice services producers globally with a special emphasis on those operating in the Southwest United States.

The Consortium of Texas Oil Producers (CTOP), has engaged FTNF to develop a study to better understand the potential for production in designated areas and help decide the best way to allocate their valuable resources.  Specifically, CTOP needs a way to predict potential energy production (and revenues generated).

FTNF is the ideal partner for this undertaking.  Our process includes enriching existing CTOP data to improve the ability to predict production outcomes.  We start with exploring the data, creating new data, and augmenting the data.  Once satisfied, we proceed to analyzing the data using contemporary statistical methods.  From there, we build predictive models to provide insight necessary to make drilling decisions.

CTOP companies no longer have to rely on hope as a strategy for finding productive wells.  With the help of FTNF Analytics, you can rely on sound science and quantifiable predictions.




# Frac Focus Data

The data related to chemicals used to enrich the data on wells comes from the chemical registry website FracFocus.com.  The FracFocus website and data is maintained by two organizations:  Interstate Oil and Gas Compact Commission and the Ground Water Protection Council.

The site maintains data on over 127,000 wells and includes not only data about which chemicals are used, but also data about groundwater protection.

## Import

The deata from FracFocus is obtained via an API call.  For the sake of reproducibility,
the API call is made outside of the code and the resulting CSV is stored in the 
program repository.

```{r fracfocus_import, cache = TRUE}
```

## Count Locations and Wellbores

```{r fracfocus_count, cache = TRUE}
```

## Aggregate Proppant and Water by wellbore

```{r fracfocus_summary, cache = TRUE}
```

## Aggregates

```{r fracfocus_aggregates, cache = TRUE}
```

# Well Features and Characteristics

```{r wellfeatures_import, cache = TRUE}
```

## Count Locations and Wellbores


```{r wellfeatures_count, cache = TRUE}
```

## Aggregates

```{r fracfocus_aggregates, cache = TRUE}
```



# Data Exploration

## Add Calculated Columns

```{r exp_welldata_calcs}
```

## Clean data

```{r exp_welldata_calcs}
```

## Summarize Well Data 

```{r exp_summary}
```

```{r exp_boxplot_fracsize}
```

```{r exp_boxplot_fracsize}
```

### Frequencies

```{r exp_freq_vintage}
```


```{r exp_freq_status}
```

# Histograms (to be added)

# Regression Model 1  (to be added)

 - Peter: Insert your model here if you have it, otherwise I will make one.

 - Well Productivity (oil.pk.bbl) vs frac.size, controlling for formation, would be a really good one.
  - Well Productivity (oil.pk.bbl) vs frac.size, controlling for formation, would be a really good one.

# Regression Model 2  (to be added)

 - Peter: Insert your model here if you have it, otherwise I will make one.

- Well Productivity (oil.pk.bbl) vs formation | perf.ll.ft + lb.ft + bbl.ft + additive.ct and controlling for formation would also be great.

# Anova to compare models



# Results of Tests



# Conclusions



# Appendix:

```{r}
sessionInfo()
```