---
title: "Measuring Well Productivity by Completion Design"
author: | 
        | Flaming  
        | Trevathan
        | Nixon  
        | Friedrich
output:
  prettydoc::html_pretty:
    keep_md: yes
    theme: leonids
    highlight: github
    toc: true
    number_sections: true
    df_print: kable
    fontsize: 11pt
    fontfamily: palatino

date: "July 24, 2018"

---

<!-- ```{r code = readLines('Setup.R'), message=FALSE, warning=FALSE} -->
<!-- ``` -->

```{r setup, echo=FALSE, results="hide"}
# run setup files, create custom functions, and read in data for analysis
source("Setup.R")
source("Functions.R")
read_chunk('DataImport-FracFocus.R')
read_chunk('DataImport-WellFeatures.R')
read_chunk('Explore_Data.R')

```


## Executive Summary

FTNF Analytics is a boutique firm specializing in statistics and data analysis for the energy sector.  The FTNF team includes experts in geology, data visualization, statistics, and predictive modeling.  Our practice services producers globally with a special emphasis on those operating in the Southwest United States.

The Consortium of Texas Oil Producers (CTOP), has engaged FTNF to develop a study to better understand the potential for production in designated areas and help decide the best way to allocate their valuable resources.  Specifically, CTOP needs a way to predict potential energy production (and revenues generated).

FTNF is the ideal partner for this undertaking.  Our process includes enriching existing CTOP data to improve the ability to predict production outcomes.  We start with exploring the data, creating new data, and augmenting the data.  Once satisfied, we proceed to analyzing the data using contemporary statistical methods.  From there, we build predictive models to provide insight necessary to make drilling decisions.

CTOP companies no longer have to rely on hope as a strategy for finding productive wells.  With the help of FTNF Analytics, you can rely on sound science and quantifiable predictions.


## Frac Focus Data Source

The Hydraulic Fracturing Disclosure and Education websites found [here](http://fracfocusdata.org) are being hosted by the Ground Water Protection Council (GWPC) and the Interstate Oil and Gas Compact Commission (IOGCC).  This website provides a central location for public and industry to communicate and relay information on the chemicals used during the process of hydraulic fracturing of oil or gas wells.  The FracFocus website provides impartial and balanced education tools to the public on the topic of hydraulic fracturing.

The GWPC and IOGCC are uniquely suited to host these websites due to their impartial nature and ties to the regulated and regulatory community. This website provides a means for industry to voluntarily supply hydraulic fracturng chemical data in a consistent and centralized location.  This open process assists both public and industry by supplying a centralized repository for the data and removing similar fragmented efforts from around the web. 

These websites provide the following:
A means to Search for voluntarily submitted chemical records by State/County, Operator and Well.
An Education and Informative site for the public and industry on Hydraulic Fracturing Chemicals and the processes used in Hydraulic Fracturing of wells.

The data related to chemicals used to enrich the data on wells comes from the chemical registry website FracFocus.com.  The FracFocus website and data is maintained by two organizations:  Interstate Oil and Gas Compact Commission and the Ground Water Protection Council.

The site maintains data on over 127,000 wells and includes not only data about which chemicals are used, but also data about groundwater protection. FracFocus has instituted a Help Desk to address any issues you may have in using the system. You can reach the Help Desk Monday-Thursday from 8 AM to 5 PM and on Friday from 8 AM to 4 PM CDT at 405-607-6808.



## Driftwood Well Data Source

The Driftwood dataset was provided by the client and can be found [here](https://github.com/la-mar/DDS_Case_Study_2/blob/master/data/deo_well_data.csv). This dataset provides a central location for wellsites to communicate and relay information on the drilling parameters used during the process of extraction of oil or gas wells. The following variables are included within the dataset:

* API - fourteen digit american petroleum institue number
* API10 - ten digit american petroleum institue number. Represents a unique hole in the ground.
* oper_alias - standardized operator (company) names
* Form_Avg - geological formation names
* PerfLL - wellbore lateral length
* FirstProd - date of first oil/gas production
* Oil_PkNorm_PerK_6mo - peak oil production within the first 6 months of first production, normalized to 1000 ft 


## FracFocus Dataset Import
### 19 of the 24 variables are selected for the study
The deata from FracFocus is obtained via an API call.  For the sake of reproducibility,
the API call is made outside of the code and the resulting CSV is stored in the 
program repository.


```{r fracfocus_import, cache = TRUE}
```

## Count Locations and Wellbores
### The FracFocus Data consists of Well Data for 3162 locations with 3163 wellbores
```{r fracfocus_count, cache = TRUE}
```

## Aggregate Proppant and Water by wellbore
### Filtering and grouping of aggregates proponents and water for each wellbore
```{r fracfocus_summary, cache = TRUE}
```

## Aggregates
### A table of FracFocus wellbore specific descriptive statistics
```{r fracfocus_aggregates, cache = TRUE}
```

## Well Features and Characteristics Dataset Import
### 7 of the 8 variables are selected for the study
```{r wellfeatures_import, cache = TRUE}
```

## Count Locations and Wellbores
### The Well Features and Characteristics consists of Well Data for 2907 locations with 2914 wellbores
```{r wellfeatures_count, cache = TRUE}
```

## Aggregates
### A table of Well Features and Characteristics wellbore specific descriptive statistics
```{r fracfocus_aggregates, cache = TRUE}
```


## Data Exploration of Well Data

Here the FracFocus and Well Features and Characteristics datasets are joined into a new dataset named welldata by the common variable api10, which represents a specific wellbore hole.


## Add Calculated Columns to Well Data
### Columns are created by the mutate function to provide 8 new variables for upcomming data analysis
```{r exp_welldata_calcs}
```

## Clean Well Data
### Here the welldata is cleaned in preparation for statistical procedures and visualization
```{r exp_welldata_calcs}
```

## Summarize Well Data 
### A table of Well Data wellbore specific descriptive statistics
```{r exp_summary}
```
## Well Data Box Plot
### A visualization of the summary statistics for Well Data
```{r exp_boxplot_fracsize}
```

```{r exp_boxplot_fracsize}
```

## Wellsite Frequencies
### A table of well frequency by respective year
```{r exp_freq_vintage}
```

## Wellsite Production Frequencies
### A table of well production by respective year
```{r exp_freq_status}
```

## Histograms (to be added)
### Plot 1
```{r}

```

### Plot 2
```{r}

```

## Regression Model 1
### Multiple Linear Regression Model for Well Productivity given Aggregate and Formation Interaction
Some of the most interesting research findings are those involving interactions among
predictor variables. Consider the welldata oil.pk.bbl measure of production. Let’s say
that you’re interested in the impact of geological formation (formavg) and aggregates (frac.size) on well production (oil.pk.bbl). You could fit a regression model that includes both predictors, along with their
interaction, as shown in the model formula below.

 - formula = lm(log(oil.pk.bbl) ~ log(frac.size) : formavg, data=welldata)  

The formula is typically written as Y ~ X1 + X2 + ... + Xk where the ~ separates the response variable log(oil.pk.bbl) on the left from the predictor variables log(frac.size) : formavg on the right, and the predictor variables are separated by + signs. The colon is used to denote an interaction between predictor variables. A prediction of y from x, z, and the interaction between x and z would be coded y ~ x + z + x:z.
```{r} 
#  Linear Regression Model for Well Productivity}
#Linear Regression Model formula = lm(log(oil.pk.bbl) ~ log(frac.size) : formavg, data=welldata)
well.production <- lm(log(oil.pk.bbl) ~ log(frac.size) : formavg, data = welldata)
```

The use of logarithmic transformations on the quantifiable variables is needed to meet the four linear regression assumptions of Normality, Independence, Linearity, and Homoscedasticity. These logarithmic transformations allow the model to meet all assumptions and our goal is to select model parameters (intercept and slopes) that minimize the difference between actual response values and those predicted by the model. Specifically, model parameters are selected to minimize the sum of squared residuals.


## Multiple Linear Regression Model for Well Productivity given Aggregate and Formation Summary
### Summary of Significant Aggregate-Formation Combinations
```{r} 
#  Linear Regression Model for Well Productivity Summary}
summary(well.production)
```
You can see from the Pr(>|t|) column that the interaction between log(frac.size) and
formavg is significant. What does this mean? A significant interaction between two
predictor variables tells you that the relationship between one predictor and the
response variable depends on the level of the other predictor. Here it means the relationship
between log(oil.pk.bbl) and log(frac.size) varies by formavg.



## Multiple Linear Regression Model for Well Productivity Plots
### Plots of Multiple Linear Regression Model for Production given all Aggregate-Formation Interactions
```{r} 
#  Linear Regression Model for Well Productivity Plots}
install.packages("effects")
library(effects)
#Plot for effect on log(Production) by Aggregate-Formation interaction 
plot(effect("frac.size:formavg", well.production,, list(wt = c(2.2, 3.2, 4.2))), multiline = TRUE)

```
You can see from this graph that as the geological formation (formavg) of the well increases in depth, the relationship between aggregate (frac.size) and well production dramatically weakens for the Spayberry Lower Shale and the Strawn formations, but there is also a gradual weakening from the Wolfcamp A formation towards the formation of Wolfcamp C, and once the well reaches the target formation of Wolfcamp C Target, there is a drastic increase in oil production through the well.

Unfortunately, fitting the model is only the first step in the analysis. Once you fit a regression model, you need to evaluate whether you’ve met the statistical assumptions underlying your approach before you can have confidence in the inferences you draw. The use of confidence intervals, overdispersion tests, and the anova test will be useful in checking if the statistical assumptions are met.

## Multiple Linear Regression Model for Well Productivity 95% CIs
### 95% CIs for Significant Aggregate-Formation Combinations
Although the summary() function in listing 8.4 describes the model, it provides no information
concerning the degree to which you’ve satisfied the statistical assumptions underlying
the model. Why is this important? Irregularities in the data or misspecifications of the relationships
between the predictors and the response variable can lead you to settle on a model that’s wildly inaccurate. On the one hand, you may conclude that a predictor and a response variable are unrelated when, in fact, they are. On the other hand, you may conclude that a predictor and a response variable are related when, in fact, they aren’t! You may also end up with a model that makes poor predictions when applied in real-world settings, with significant and unnecessary error. Let’s look at the output from the confint() function applied to the multiple regression problem under study.
```{r} 
#  Linear Regression Model for Well Productivity 95% CIs}
#95% CI for Linear Regression Model
confint(well.production)
```
The results suggest that you can be 95% confident that the intervals that DO NOT contain 0 give the true change in well production rate for their respected % change in aggregate:formation interaction rate. Remember that, because the confidence intervals for the DEAN, SPBY_L_SILT, and STRAWN formations contain 0, you can conclude that a change in their rate is unrelated to the well production rate, holding the other variables constant. But your faith in these results is only as strong as the evidence you have that your data satisfies the statistical assumptions underlying the model.

A set of techniques called regression diagnostics provides the necessary tools for evaluating
the appropriateness of the regression model and can help you to uncover and
correct problems. We’ll start with a standard approach that uses functions that come
with R’s base installation. Then we’ll look at newer, improved methods available
through the car package.


## Overdispersion Check for Linear Model
### Rebuild Model if considerally larger than 1
```{r} 
#  Check222 for Overdispersion of Logistic Regression Model}
#Residual deviance divided by residual degrees of freedom is used to detect overdispersion in a binomial model, if considerably larger than 1, you have evidence of overdispersion
overdispersion_test1 <- deviance(well.production)/df.residual(well.production)
overdispersion_test1
```


## Results of Linear Regression Model for Well Productivity given all Aggregate-Formation Combinations
```{r} 
# Results for Anova Test}
#Comparison of the One-Way Anova Tests for Linear Regression Model
anova(well.production)

```


## Conclusions
From the p-values for the Linear Regression Model coefficients (last column) from the anova() output, you can see that the explanitory variable frac.size given formavg makes a significant contribution to the equation (you can reject the hypothesis that the parameters are 0). With that being evident, there is no need to fit a second reduced model that may fit the data as well. The log-log Linear Regression Model shows that each regression coefficient in the model is statistically significant (p < .05). The anova() function for linear models uses a F-test resulting in an (F-value = 5.9052) and (Pr(>F) < 0.0001). The nonsignificant  F-value (p < 0.0001) suggests that the model fits well, reinforcing your belief that frac.size given formavg adds significantly to the model above and therefore, you can base your interpretations on the resulting linear model. 


## Regression Model 2  
### Logistic Regression Model for Successfully Drilling the Target Formation  
```{r} 
#  fracfocus_registry}
library(readr)
fracfocus_registry <- read_csv("~/Documents/GitHub/DDS_Case_Study_2/data/fracfocus_registry.csv",
col_types = cols(API10 = col_number(),
APINumber = col_skip(), ClaimantCompany = col_skip(),
CountyNumber = col_number(), DTMOD = col_skip(),
FFVersion = col_number(), FederalWell = col_logical(),
IndianWell = col_logical(), IngredientComment = col_skip(),
IngredientMSDS = col_logical(), IsWater = col_skip(),
JobEndDate = col_date(format = "%m/%d/%Y"),
JobStartDate = col_date(format = "%m/%d/%Y"),
Latitude = col_number(), Longitude = col_number(),
MassIngredient = col_number(), PercentHFJob = col_skip(),
PercentHighAdditive = col_skip(),
Source = col_skip(), StateNumber = col_skip(),
Supplier = col_skip(), SystemApproach = col_skip(),
TVD = col_number(), TotalBaseNonWaterVolume = col_number(),
TotalBaseWaterVolume = col_number(),
TradeName = col_skip(), pKey = col_number(),
pKeyDisclosure = col_number(), pKeyPurpose = col_skip(),
pKeyRegistryUpload = col_skip()),
na = "0")
View(fracfocus_registry)
str(fracfocus_registry)
TVD <- as.data.frame.numeric(fracfocus_registry$TVD)
TBWV <- as.data.frame.numeric(fracfocus_registry$TotalBaseWaterVolume)
TBNonWV <- as.data.frame.numeric(fracfocus_registry$TotalBaseNonWaterVolume)

fracfocus_registry_clean <- drop_na(fracfocus_registry)

summary(fracfocus_registry_clean)
```

```{r} 
#  deo_well_data}
deo_well_data <- read_csv("~/Documents/GitHub/DDS_Case_Study_2/data/deo_well_data.csv")
deo_well_data["PerfLL"] <- as.numeric(deo_well_data$PerfLL)
deo_well_data["FirstProd"] <- as.numeric(deo_well_data$FirstProd)
deo_well_data["Oil_PkNorm_Perk_6mo"] <- as.numeric(deo_well_data$Oil_PkNorm_Perk_6mo)

str(deo_well_data)

deo_well_data_clean <- drop_na(deo_well_data)

summary(deo_well_data_clean)
```

```{r} 
#  Permian Basin Well Data Selection}
library(tidyverse)
Basin_Data <- merge.data.frame(deo_well_data, fracfocus_registry, by = intersect(x="API10", y="API10"))
summary(Basin_Data)
```

## Data Preperation for Logistic Regression Model
Logistic regression is applied to situations in which the response variable is dichotomous (0 or 1). The model assumes that Y follows a binomial distribution and that you can fit a linear model of the form where π = μY is the conditional mean of Y (that is, the probability that Y = 1 given a set of X values), (π/1 – π) is the odds that Y = 1, and log(π/1 – π) is the log odds, or logit. In this case, log(π/1 – π) is the link function, the probability distribution is binomial, and the logistic regression model can be fit using glm(Y~X1+X2+X3, family=binomial(link="logit"), data=mydata)

```{r} 
#  Logistic Model for Basin_Data}
library(AER)
#Creating Logistic Model binary outcome for TVD greater than 8782ft set as yes, and less than 8782 as no
Basin_Data$ynTVD[Basin_Data$TVD <= 8394] <- 0
Basin_Data$ynTVD[Basin_Data$TVD >= 8394] <- 1
#Create a dichotomous factor to be used as the outcome variable in a logistic regression model
Basin_Data$ynTVD <- factor(Basin_Data$ynTVD, levels=c(0,1), labels=c("No","Yes"))
#View table of output for new predictor variable ynTVD
table(Basin_Data$ynTVD)
#Selecting Basin variables for Logistic Regression and creating new dataset for analysis
Basin_Model_Data <- Basin_Data[, c(4, 5, 8, 15, 16, 17, 18, 22, 32)]
head(Basin_Model_Data)
#Descriptive Statistics for Model variables
summary(Basin_Model_Data)
```
The target depth is 8394ft. for the WFMP_C_TARGET Formation. 

The glm() function allows you to fit a number of popular models, including logistic regression, Poisson regression, and survival analysis (not considered here). You can demonstrate this for the first two models as follows. Assume that you have a single response variable (Y), three predictor variables (X1, X2, X3), and a data frame (mydata) containing the data.

Logistic regression is useful when you’re predicting a binary outcome from a set of continuous and/or categorical predictor variables. To demonstrate this, let’s explore the well data for total vertical depth of target formation contained in the data frame under study. Be sure to download and install the package (using install.packages("AER")) before first use.
```{r} 
#  Full Logistic Regression Model}
#Full Model formula = ynTVD ~ form_avg + Status + Oil_PkNorm_Perk_6mo + Latitude + Longitude + Projection + CountyName, family = binomial(), data = Basin_Model_Data
fit.full.basin <- glm(ynTVD ~ form_avg + Status + Oil_PkNorm_Perk_6mo + Latitude + Longitude + Projection + CountyName, data=Basin_Model_Data, family=binomial(link='logit'))
```


```{r} 
#  Full Logistic Regression Model Summary}
#Descriptive Statistics for the Full Logistic Regression Model
summary(fit.full.basin)
#There are three predictors ("SPBY_L_SILT", "SPBY_M", and "WFMP_D") and one predictor variable ("Status") that are not statistically significant
```
From the p-values for the regression coefficients (last column), you can see that there are three predictors ("SPBY_L_SILT", "SPBY_M", and "WFMP_D") and one predictor variable ("Status") that are not statistically significant (p-values > 0.05) and they may not make a significant contribution to the equation (you can’t reject the hypothesis that the parameters are 0). Let’s fit a second equation without them and test whether this reduced model fits the data as well:


```{r Full Logistic Regression Model Plots}
influencePlot(fit.full.basin)
```
The horizontal axis is the leverage, the vertical axis is the studentized residual, and the plotted symbol is proportional to the Cook’s distance. Diagnostic plots tend to be most helpful when the response variable takes on many values. When the response variable can only take on a limited number of values (for example, logistic regression), the utility of these plots is decreased.You can see from these graphs that as the geological formation (form_avg) of the well increases in depth, the relationship between those significant predictors explane the response of target formation depth throughout the well drilling process.

```{r} 
#  Full Logistic Regression Model 95% CIs}
#95% CI for Full Logistic Model
exp(confint(fit.full.basin))
```
The results suggest that you can be 95% confident that the intervals that DO NOT contain 0 predict the true change in well production rate for their respected % change in aggregate:formation interaction rate. Remember that, because the confidence intervals for the DEAN, SPBY_L_SILT, and STRAWN formations contain 0, you can conclude that a change in their rate is unrelated to the well production rate, holding the other variables constant. But your faith in these results is only as strong as the evidence you have that your data satisfies the statistical assumptions underlying the model.

```{r} 
#  Reduced Logistic Regression Model}
#Reduced Logistic Model formula = ynTVD ~ form_avg + Oil_PkNorm_Perk_6mo + Projection, family = binomial(), data = Basin_Model_Data
fit.reduced.basin <- glm(ynTVD ~ form_avg + Oil_PkNorm_Perk_6mo + Projection, data=Basin_Model_Data, family=binomial(link='logit'))
```


```{r} 
#  Reduced Logistic Regression Model Summary}
#Descriptive Statistics for the Reduced Logistic Regression Model
summary(fit.reduced.basin)
```


```{r} 
#  Reduced Logistic Regression Model Plots}
#Plot of reduced Logistic Model
influencePlot(fit.reduced.basin)
```
The horizontal axis is the leverage, the vertical axis is the studentized residual, and the plotted symbol is proportional to the Cook’s distance. Diagnostic plots tend to be most helpful when the response variable takes on many values. When the response variable can only take on a limited number of values (for example, logistic regression), the utility of these plots is decreased.You can see from these graphs that as the geological formation (form_avg) of the well increases in depth, the relationship between those significant predictors explane the response of target formation depth throughout the well drilling process.

```{r} 
# Reduced Logistic Regression Model 95% CIs}
#95% CI for Reduced Logistic Model
exp(confint(fit.reduced.basin))
```



```{r} 
#  Check for Overdispersion of Logistic Regression Model}
#Residual deviance divided by residual degrees of freedom is used to detect overdispersion in a binomial model, if considerably larger than 1, you have evidence of overdispersion
overdispersion_test <- deviance(fit.reduced.basin)/df.residual(fit.reduced.basin)
overdispersion_test
```


```{r} 
#  Hypothesis Test for H0 = 1 and Ha != 1}
#First instance
fit <- glm(ynTVD ~ form_avg + Oil_PkNorm_Perk_6mo + CountyName + Latitude + Longitude + Projection, data=Basin_Model_Data, family=binomial())
#Second instance
fit.od <- glm(ynTVD ~ form_avg + Oil_PkNorm_Perk_6mo + Projection, data=Basin_Model_Data, family = quasibinomial())

#If the p-value is small (say, less than 0.05), you'd reject the null hypothesis
pchisq(summary(fit.od)$dispersion * fit$df.residual, fit$df.residual, lower = F)
#The resulting p-value is clearly significant (p < 0.05), strengthening the belief that overdispersion isn't a problem
```
The resulting p-value is clearly significant (p < 0.05), strengthening the belief that overdispersion isn't a problem

Each regression coefficient in the reduced model is statistically significant (p < .05). Because the two models are nested (fit.reduced is a subset of fit.full), you can use the anova() function to compare them. For generalized linear models, you’ll want a chi-square version of this test.


## Two-Way Anova for Reduced Logistic Regression Model
### For the two generalized linear models, likelihood-ratio chisquare, Wald chisquare, or F-tests are calculated
```{r} 
#  Results for Anova Chi-Suared Test}
#Comparison of the Two-Way Anova Tests for Full and Reduced Logistic Regression Model
anova(fit.reduced.basin, fit.full.basin, test = "Chisq")

```
The nonsignificant chi-square value (Chi < 0.0001) suggests that the reduced model with three predictors fits as well as the full model with six predictors, reinforcing your belief that status, latitude, longitude, and countyname don’t add significantly to the prediction above and beyond the other variables in the equation. Therefore, you can
base your interpretations on the simpler model.

## Conclusions
From the p-values for the Full Logistic Regression Model coefficients (last column) of the anova.full model output, you can see that the predictor variable Status may not make a significant contribution to the equation (you can’t reject the hypothesis that the parameters are 0). With that being evident, there is a need to fit a second equation without the Status variable and test whether this reduced model fits the data as well. The Reduced Logistic Regression Model shows that each regression coefficient in the reduced model is statistically significant (p < .05). Because the two models are nested (fit.reduced is a subset of fit.full), you can use the anova() function to compare them. For generalized linear models, you’ll want a chi-square version of this test. The nonsignificant chi-square value (Chi < 0.0001) suggests that the reduced model with three predictors fits as well as the full model with seven predictors, reinforcing your belief that Status, CountyName, Latitude, and Longitude don’t add significantly to the prediction above and beyond the other variables in the equation. Therefore, you can base your interpretations on the simpler model. 


# Appendix:

```{r}
sessionInfo()
```