---
title: "qn_initial_EDA"
author: "Q Nixon"
date: "7/19/2018"
output: html_document
---

## Get the data

I'm switching between working directories to pull the data from the main repository
and store everything in a subfolder.  Run the file "qn_initial_EDA.R" first before
proceeding with the EDA in this markdown file.

```{r, echo=TRUE}
setwd("~/Documents/@Grad School/@SMU/@Term 1 Courses/Doing Data Science/Week11/DDS_Case_Study_2/data")
well_data <- read.csv("deo_well_data.csv", stringsAsFactors = FALSE)
registry_data <- read.csv("fracfocus_registry_texas.csv", stringsAsFactors = FALSE)
well_data$FirstProd <- as.Date(well_data$FirstProd, "%m/%d/%Y")
well_data$Oil_PkNorm_PerK_6mo <- as.numeric(well_data$Oil_PkNorm_PerK_6mo)
setwd("~/Documents/@Grad School/@SMU/@Term 1 Courses/Doing Data Science/Week11/DDS_Case_Study_2/qn_code_and_plots")
```

## Initial plots

Below are various plots looking at different variables in the data files.  

```{r, echo=TRUE}
peak_performance_wells <- subset(well_data, !is.na(well_data$Oil_PkNorm_PerK_6mo))
boxplot(log(peak_performance_wells$Oil_PkNorm_PerK_6mo))
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
